---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# fastmart

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/fastmart)](https://CRAN.R-project.org/package=fastmart)
[![R-CMD-check](https://github.com/selkamand/fastmart/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/selkamand/fastmart/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

The goal of fastmart is to speed up a subset of biomaRt queries by preparing local, properly indexed sqlite databases.


## Installation

You can install the development version of fastmart like so:

``` r
install.packages("fastmart", type = "source", repos = "https://github.com/selkamand/fastmart")
```

## Quick start

The first time you use this package,  you will need to **initialise** fastmart and cache the tables we need.
This can be done as follows

```{r example, eval = FALSE}
library(fastmart)

# Initialise fastmart (creates a ~/.fastmart directory key databases will be stored)
fastmart_init()

# Cache the key tables
fastmart_cache_tables(GRCh = "37") 
```

You can now use any of the supported quick lookups

## Available Functionality

### Convert Gene IDs

Convert HGNC Symbols to ENSEMBL gene IDs

```{r hgnc_to_ensembl, eval = FALSE}
# Convert HGNC to Ensemble IDs
fastmart_hgnc_to_ensembl('TBCE', chrom='1', start='235530675', end='235612283', GRCh = "37")
  # > ENSG00000116957.8
```

